---
title: "attendersCategorisation"
author: "Dave Lovell"
date: "26/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library("dplyr")
library("here")
library("magrittr")
library("forcats")
library("purrr")
library("stringr")
library("lubridate")
```
## Some Old Code

```{r someOldCode}
##  black-box reverse engineering the church background categories
## DL 190912

fileInO <- "O:\\WCC\\Learning and Development\\Research\\PROJECT - Portsmouth 2019 onwards\\Strand 1 - Leaders and Attenders Surveys\\Data\\2018 Data Processing\\rdats\\attenders05-190911 fixing ages.rdat"

if(file.exists(fileInO))
{
  oldAttenders = readRDS(fileInO)
  file.copy(fileInO, here::here("data//oldAttenders.rdat"), overwrite = TRUE)
} else {
  message("Acquiring old attenders data locally - O Drive not connected")
  oldAttenders = readRDS(here::here("data/oldAttenders.rdat"))
}

glimpse(oldAttenders)

newAttenders <- readRDS(here::here("data/2020attenders.rdat"))
## Also load the 2020 data
```

For this exercise, we'll only need some of the data

```{r dropBareCols}
# Little bit of implicit renaming for tranche compatibility
# oldAttenders %<>% select(responseType, ageGroup, regularity, transferStatus, transferReason, transferOther, christian, class1, class2
#                          becameHere = becameChristianHere,
#                          statusChild = childStatus,
#                          statusTeen = teenStatus,
#                          statusYoung = youngStatus,
#                          statusAdult = adultStatus)

## Mutate factor levels for consistency
oldAttenders  %<>% mutate(churchBackground = transferStatus %>%
                           fct_recode("churched" = "fromChurch", 
                                       "deChurched" = "fromBreak",
                                       "blending" = "twoChurches", 
                                       "nonChurched" = "firstChurch"))
oldAttenders %<>% select(-transferStatus)
oldAttenders %<>% rename("statusChild" = "childStatus",
                         "statusTeen" = "teenStatus", 
                         "statusYoung" = "youngStatus",
                         "statusAdult" = "adultStatus")
oldAttenders %<>% mutate(class1 = 
                           fct_recode(class1,
                                      "complexDe" = "copmlexDe"))
oldAttenders %<>% mutate(ageGroup = 
                           fct_recode(ageGroup,
                                      "0-4" = "under5"))

### Get rid of children
oldAttendersAdult <- oldAttenders %>%  
  filter(ageGroup != "0-4", responseType == "adult")

### Get rid of responses which can be assigned class1: broughtUp

oldAttendersAdult %<>% 
  filter(!(ageGroup %in% c("0-4", "5-10", "11-15") & # under 16 years old
           regularity %in% c("almostAlways", "halfTime", "fewTimes") & # attend at least a few times a year
           transferReason == "fromYoung")) # say they grew up here

### Get rid of everyone who's certifiably 'churchedChurched'

oldAttendersAdult %<>%
  filter(!(statusChild %in% c("church", NA) &
         statusTeen %in% c("church", NA) &
         statusYoung %in% c("church", NA) &
           statusAdult %in% c("church", NA) &
           churchBackground %in% c("blending", "churched")))

oldAttendersAdult %<>%
  select(churchBackground,
         statusChild,
         statusTeen,
         statusYoung,
         statusAdult,
         class1)
```

And now a special function which gets the 'funky' church histories removed

```{r removeFunkyHistories, warning = FALSE}
## This function looks at the life-stage data and flags anything which has an NA _before_ the latest entry (i.e. incomplete data)
determineFunkiness <- function(someData){
  someData %<>% 
    mutate(isFunky = 
             pmap_lgl(list(statusChild, statusTeen, statusYoung, statusAdult),
             ~ {
               if(all(is.na(c(..1, ..2, ..3, ..4)))) TRUE else { 
               c(..1, ..2, ..3, ..4) %>% 
                 (negate(is.na)) %>% 
                 which %>% 
                 {!(all(. == 1:max(.)))}}}))
  return(someData)
}

oldAttendersAdult %<>% determineFunkiness
oldAttendersAdult %<>% filter(!isFunky)
oldAttendersAdult %<>% select(-isFunky)

```

Okay this might get confusing: first, find all existing permutations. Then, find any permutations which **always** result in the same `class1` designation

```{r permutations}

### Function for determining mode
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

permutations <- 
  oldAttendersAdult %>%
  group_by_at(vars(-class1)) %>% 
  mutate(mode = Mode(class1)) %>% 
  group_by_all %>%
  summarise(n())



permutations %>%
  group_by_at(vars(-`n()`)) %>%
  summarise(total = sum(`n()`)) %>% 
  group_by_at(vars(-total, -class1)) %>% 
  summarise(biggest = max(total),
            percentage = max(total) / sum(total) * 100) -> decisionSummary

  # summarise(class1sAssigned = length(unique(class1)), `n()` = sum(`n()`)) %>% 
  # arrange(class1sAssigned, -`n()`) %>%
  # filter(class1sAssigned == 1) %>% 
  # select(- class1sAssigned) %>% 
  # left_join(oldAttenders, by = c("churchBackground", "statusChild", "statusTeen", "statusYoung", "statusAdult")) %>% 
  # distinct %>% 
  # arrange(-`n()`, class1) %>% 
  # View
```

### Try to Understand Contentious Cases

```{r analyseContentious}
### Get the notes from contentious decisions
decisionSummary %>% 
  filter(percentage < 65) %>% 
  select(-mode, - biggest, -percentage) %>% 
  left_join(oldAttenders %>% 
              filter(ageGroup != "under5", responseType == "adult") %>% 
              select(-responseRef, - responseId, - responseType, -gender, - class2 , - university, - ethnicity, - plantName, - x10Years),
            by = c("churchBackground", "statusChild", "statusTeen", "statusYoung", "statusAdult")) -> contentiousReasons
```
Complex cases are, unsurprisingly, harder to decode.

```{r makeHistoryListColumn}
makeHistoryList <- 
  function(someData){
    someData %>% 
      rowwise %>% 
      mutate(history = 
               list(c(as.character(statusChild),
                      as.character(statusTeen),
                      as.character(statusYoung),
                      as.character(statusAdult)))) %>% 
      ungroup -> someData
    return(someData)
  }

oldAttendersAdult %<>% makeHistoryList

oldAttendersAdult %>% 
  rowwise %>% 
  filter(history %>% 
           {
             sum(is.na(unlist(.)) | unlist(.) %in% "(Missing)") < 3
           }) %>% 
  filter(history %>%
           {
             unlist(.) %>% 
               {.[(max(which(!is.na(.)))- 1) : max(which(!is.na(.)))]} %>% 
               {all(. %in% "noChurch")}
           }) 
```
### Categorising 2020 attenders

```{r assignPlantIds}
### Assign plantIds

newAttenders %<>% mutate(ageGroup = 
                           fct_recode(ageGroup,
                                      "0-4" = "0-5",
                                      "5-10" = "6-10"))

### Fix a bad error. Not sure why this happened:
### Shouldn't be possible to save adult surveys as child, but also shouldn't be possible
### to put wrong age on child surveys. I've checked all of these - they're all in fact adult surveys.

newAttenders %<>% rowwise

newAttenders %<>% 
  mutate(responseType = 
           if(!(ageGroup %in% c("0-4", "5-10")) |
              (
                ageGroup %in% c("(Missing)", NA) &
               (
                 !is.na(statusChild) |
                 !is.na(statusTeen) |
                 !is.na(statusYoung)|
                 !is.na(statusAdult)
               )
               )) "adult"
                else responseType)
             
newAttenders %>% 
  rowwise() %>% 
  mutate(plantId = stringr::str_sub(digest::sha1(location), 1, 10)) -> newAttenders
```
Get 'year started' and 'year returned'

```{r yearStartedAndYearReturned}
newAttenders %<>%
  mutate(yearStarted = 
           if(statusAdult %in% "notPart") statusAdultDetail
         else NA)

newAttenders %<>% 
  mutate(yearReturned =
           if(statusAdult %in% "stoppedHere") statusAdultDetail
         else NA)

### Clean those things

newAttenders %<>% 
  mutate(across(all_of(c("yearStarted", "yearReturned")),
                ~ replace(., . == "", NA) %>% 
                  str_remove_all("[^\\d[:space:]]") %>% 
                  str_extract("[^[space]]{4}$") %>% 
                  as.numeric)) 
```

```{r assignClass1ForAdults}
### class1:broughtUp

newAttenders %<>% determineFunkiness()

newAttenders %<>% 
  mutate(class1 = 
           if(!isFunky &
              responseType == "adult" &
              regularity %in% c("almostAlways", "halfTheTime", "fewTimesAYear") &
              transferReason == "grewUp" & 
              ageGroup %in% c("0-4", "5-9", "10-15")) "broughtUp"
         else NA)

### class1:churchedChurched
newAttenders %<>%
  mutate(class1 = 
           if(!is.na(class1)) class1 
         else if(!isFunky &
                 responseType == "adult" &
                 statusChild %in% c("wasPart", NA) &
                 statusTeen %in% c("wasPart", NA) &
                 statusYoung %in% c("wasPart", NA) &
                 statusAdult %in% c("wasPart", NA) &
                 churchBackground %in% c("churched", "blending")) "churchedChurched"
         else NA
           )

### class1:simpleDe
newAttenders %<>%
  mutate(class1 = 
           if(!is.na(class1)) class1 
         else if(!isFunky &
                 responseType == "adult" &
                 churchBackground %in% "deChurched") "simpleDe"
         else NA
           )

### class1:simpleNon
newAttenders %<>% 
  mutate(class1 = 
           if(!is.na(class1)) class1
         else if(!isFunky &
                 responseType == "adult" &
                 churchBackground %in% "nonChurched" &
                 statusChild %in% c("notPart", NA) &
                 statusTeen %in% c("notPart", NA) &
                 statusYoung %in% c("notPart", NA) &
                 statusAdult %in% c("notPart", NA) &
                 (is.na(yearStarted) | yearStarted > 2000)) "simpleNon" 
         else NA)
```

```{r getStartDates}
allLeaders <- readRDS("data/allLeaders.rdat")

newAttenders %<>%
  mutate(fxcName =
           str_replace(fxcName, "Messy Church (?=[^-])",
                       "Messy Church - "))

## Where names match, put ID on attenders
left_join(newAttenders, allLeaders %>% select(fxc, fxcId) %>% distinct,
          by = c("fxcName" = "fxc")) -> newAttenders
```

Attach IDs manuall for others:

```{r manualIDAttachment}
namesAndIds <-
  as_tibble(
    matrix(
      c(
      c("Cafe Connect", "f63df641c9b" ),
      c("Chatty Church Southsea St Luke", "f56f56b76de"),
      c("Encompass", "f99bc8532cf"),
      c("Little Fishes Fareham HT" , "fcdee06f8f1"),
      c("Messy Church - Copnor", "f97c6ff218e"),
      c("Messy Church - Hook w Warsash", "f0bf8df288c"),
      c("Messy Church - Southsea St Lukes", "f19bddeae21"), 
      c("Older Persons Bluewater", "fcae33a6cd6"), 
      c("St Margarets", "ffec32c0180"),
      c("Tea Service Sarisbury", "f7e8d28c572"), 
      c("Tots in Tow Southsea St Lukes", "fcf545587ed"), 
      c("Tots Praise Hook", "f98b9011146")
      ),
      nrow = 2
    ) %>% 
      t
  )

names(namesAndIds) <- c("fxc", "fxcId")

# Will this work?

newAttenders %>% 
  rowwise %>% 
  mutate(fxcId =
           replace(fxcId,
                   is.na(fxcId) & (fxcName %in% namesAndIds[["fxc"]]),
                   namesAndIds[["fxcId"]][ namesAndIds[["fxc"]] %in% fxcName ])) %>% 
  select(-plantId) -> newAttenders

```

Now Join the start dates

```{r joinStartDates}

## Finally putting start dates on these attenders!
allLeaders %>% 
  filter(!is.na(info_startDate)) %>% 
  group_by(fxcId) %>%
  mutate(is_min = as.numeric(info_startDate) == min(as.numeric(info_startDate))) %>% 
  filter(is_min) %>% 
  select(fxcId, info_startDate) %>% 
  distinct %>% 
  right_join(newAttenders) -> newAttenders
```

## Did Returners Return to FxC?

It's guess work, but we inform guesses by comparing start date of fxC with date of return.

```{r makeCameBackHereCol}
newAttenders %<>% mutate(cameBackHere = coalesce(yearStarted, yearReturned) > as.numeric( str_sub(info_startDate, 1 , 4))) 
```

### Classify Blenders

If someone's blending and the church they returned to is (probably) this fxC, we'll list them as simpleDe. If they returned 'to Church' before this fxC began, they must be complex DeChurched. 

```{r classifyBlenders}
newAttenders %>%
  rowwise %>% 
  mutate(class1 = 
           {
             if (!is.na(class1)) class1
             else if ((statusChild %in% "stoppedHere" |
                     statusTeen %in% "stoppedHere" |
                     statusYoung %in% "stoppedHere" |
                     statusAdult %in% "stoppedHere") &
                     churchBackground == "blending"){
               if (cameBackHere %in% TRUE) "simpleDe"
               else if (cameBackHere %in% FALSE) "complexDe"
               else NA
             } else NA }) -> newAttenders

newAttenders %<>% makeHistoryList

newAttenders %<>% rowwise


#### Can't remember what I was trying to do, unsure if it's something that ought to be done.
newAttenders %>% 
  mutate(class1 = 
           if (churchBackground == "blending" !isFunky & sum(is.na(unlist(history))) < 3){
             if (all(unlist(history)[ (max(which(!is.na(history))) - 1) : max(which(!is.na(history))) ]) %in% "wasPart") 
           })
```